\section{Introduction}\label{sec:intro}

\Objectoriented languages strive to offer great code reuse.
They couple flexibility and rigour, expressive power and
modular reasoning.  Two main OO models emerged to this end:
\prototypebased (PB)~\cite{Ungar87self} and \classbased languages such as
Java, C\# or Scala.  In \prototypebased
languages objects inherit from other objects. Thus objects own
both behaviour and state (and objects are all you have).
In \classbased languages an object is an instance of a specific class,
and classes inherit from other classes.  Objects own state,
while classes contain behaviour and the structure of the state.

Regardless of the OO model, inheritance is a key mechanism is OO 
languages. Inheritance provides a modularization mechanism, which 
is used to reuse implementations from inherited classes/objects. 
Unfortunately, as widely acknowledged in the literature~\cite{scharli03traits,Sak89dis,bracha90mixin,malayeri2009cz}
multiple inheritance, especially when combined with state, has several
tricky issues, including several variants of the famous 
\emph{diamond problem}~\cite{bracha90mixin,Sak89dis}. Many of the problems related to 
inheritance arise from the direct use of \emph{fields} to model state. 
Inheriting two fields with the same name raises 
the question of whether the two fields should be kept, or only one 
field should exist. Initialization 
of the fields is also problematic, since initialization code may be inherited from 
multiple parents. Finally, an additional problem with mutable fields is that their 
type cannot be type-refined in extensions, which can cause modularity problems
~\cite{wadler98expression,eptrivially}.

To address those limitations, this paper presents a third alternative
OO model called \emph{\interfacebased} \objectoriented programming
languages (IB), where objects implement interfaces directly and fields
are not directly supported. In IB interfaces own the implementation for the
behaviour, which is structurally defined in their
interface. Programmers do not define objects directly but delegate the
task to \emph{object interfaces}, whose role is similar to
non-abstract classes in \classbased \objectoriented programming
languages (CB). Objects are instantiated by static factory methods in
object interfaces.

Due to the absence of fields, a key challenge in IB lies in how to model state, which is
fundamental to have stateful objects. All abstract operations in an
object interface are interpreted as \emph{abstract state
  operations}. The abstract state operations include various common
utility methods (such as getters and setters, or clone-like
methods). Objects are only responsible to define the ultimate
behaviour of a method. % and, for example, if such a method is just a setter.
Anything related to state is completely contained in the
instances and does not leak into the inheritance logic.  In CB, the structure of the state is fixed and can only grow
by inheritance.  In contrast, in IB the state is never
fixed, and methods such as abstract setters and getters
can always receive an explicit implementation down in the inheritance
chain, improving \textbf{modularity and flexibility}.  That is, the
concept of abstract state is more fluid.

Object interfaces provide support for automatic type-refinement.  In
contrast, in CB special care and verbose explicit type-refinement are
required to produce code that deals with subtyping adequately. We
believe that such verbosity hindered and slowed down the discovery of
useful programming patterns involving type-refinement. Our previous work~\cite{eptrivially} 
on the Expression Problem~\cite{wadler98expression} in Java-like languages shows how
easy it is to solve the problem using only type-refinement. However it
took nearly 20 years since the formulation of the problem for that
solution to be presented in the literature. In IB, due to its emphasis
on type-refinement, that solution should have been more obvious.

One advantage of abstract state operations and
type-refinement is that it allows a new approach to
\emph{type-safe covariant mutable state}. That is, in IB,
it is possible to type-refine \emph{mutable} state in subtypes. This is
typically forbidden in CB: it is widely known that \emph{naive} type-refinement of
mutable fields is not type-safe. Although covariant refinement of mutable
fields is supported by some type systems~\cite{bruce98astatically,bruce1994paradigmatic,ernst06virtual,Saito2013933}, this requires
significant complexity and restrictions to ensure that all uses of
covariant state are indeed type-safe.

\begin{comment}
\marcoT{%

In this paper we show how to support type-safe
\textbf{field removal},
\textbf{field type refinement} allowing a kind of covariant setters refinement,
and \textbf{multiple inheritance}.}
\end{comment}

\begin{comment}
Another advantage of IB is the support for
multiple inheritance. The literature shows how
easy it is to combine multiple sources of pure behaviour
using traits~\cite{scharli03traits}. In Java
multiple \emph{interface} inheritance has been supported since
inception, and Java 8 default methods~\cite{goetz12fdefenders} bring some of the
advantages of traits to Java. The literature~\cite{Sak89dis,bracha90mixin,malayeri2009cz}
is also rich on how hard it is to combine multiple sources
of behaviour \textbf{and} state with multiple \emph{implementation}
inheritance of classes. In IB there is only multiple
interface inheritance, yet programmers can still use state via
abstract state operations. IB enables powerful
idioms using multiple inheritance and state.
\end{comment}

%Traits or Java8 interfaces still assume a CB model:
%But traits still require classes, which
%are responsible for object construction and adding state.

%To retain the easiness and modularity of combining multiple sources of
%pure behaviour, in IB state is just a special kind of behaviour
%implementation.
%\marco{NO NO!
%the advantage is not just multiple composition
%(by the way, that is the term you are searching for, not multiple inheritance)
%By not having state reified in the ideal/platonic level, it is not fixed
%what is state and what is not, thus you can (model an encoding of) REMOVE fields and you can refine
%fields.}
% Objects are the only responsibles to define the
%ultimate behaviour of a method and, for example, if such method is
%just a setter. Anything related to state is completely contained in
%the instances and do not leak in the interfaces.

IB could be explained by defining a novel language, with new syntax
and semantics. However, this would have a steep learning curve.  We
take a different approach instead. For the sake of providing a more
accessible explanation, we will embed our ideas directly into Java.
Our IB embedding relies on the
new features of Java 8: interface \emph{static methods} and
\emph{default methods}, which allow interfaces to have method
implementations. In the context of Java, what we propose is a programming
style, where we never use classes (more precisely, we never
  use the \Q@class@ keyword).  We call this restricted version of
Java \emph{Classless Java}. 
\begin{comment}
\textcolor{red}{Classless Java provides us
with a way to address the famous diamond problem~\cite{bracha90mixin,Sak89dis}
based on Java 8's interface model: interfaces can offer different implementations
using the default methods, whereas conflicts are supposed to be explicitly resolved,
otherwise the program will be rejected by the compiler. An interface that inherits
a few default implementations of a method from its supertypes can define a new implemention
to hide the conflicts among them, or use the \Q@super@ keyword to specify the one it wants. Examples
and more discussion will be presented afterwards.}
\end{comment}

%Still, the design is quite conservative and appears to be quite limited
% in its current form to model advanced forms of multiple inheritance.
%Indeed, our own personal experience of combining default methods
%and multiple interface inheritance in Java to achieve multiple implementation
%inheritance is that many workarounds and boilerplate code are needed.
%In particular, we encountered difficulties because:

%\begin{itemize}
%
%\item {\em Interfaces have no constructors.} As a result, classes are
%still required to create objects, leading to substantial boilerplate
%code for initialization.
%
%\item {\em Interfaces do not have state.} This creates a tension between
% using multiple inheritance and having state. Using setter and
%  getter methods is a way out of this tension, but this workaround
%  requires tedious boilerplate classes that later implement those
%  methods.
%
%\item {\em Useful, general purpose methods require special care in
%  the presence of subtyping.} Methods such as
%  \emph{fluent} setters~\cite{fowler2005fluentinterface}, not only require access to the
%  internal state of an object, but they also require their return types to be
%  refined in subtypes.
%
%\end{itemize}

%\noindent Clearly, a way around those difficulties would be to change
%Java and just remove these limitations. Scala's own notion of
%traits~\cite{scala-overview}, for example, allows state in traits. Of
%course adding state (and other features) to interfaces would
%complicate the language and require changes to the compiler, and this
%would go beyond the goals of Java 8 development team.

%This paper takes a different approach. Rather than trying to get
%around the difficulties by changing the language in fundamental ways,
%we show that, with a simple language feature, default methods and
%interface inheritance are in fact very expressive.





%With object interfaces, many Java programs can be built
%without using a single class!

Using Java annotation processors, we produce an implementation of
Classless Java, which allows us to stick to pure Java 8. 
%By annotating
% the interfaces that represent object interfaces with \mixin, code for interface instantiation and
%type refinement can be automatically generated. Such code should not be needed in the first place in
%a real IB language, and the annotation processor allows us to
%transparently hide it from Java programmers. 
The implementation works
by performing AST rewriting, allowing most existing Java tools (such as
IDEs) to work out-of-the-box with our implementation. Moreover, the
implementation blends Java's conventional CB style and IB smoothly.
We apply object interfaces to several interesting
Java programs and conduct various case studies. Finally, we also
discuss the behaviour of Classless Java and
its properties.

%Since Java was not designed to be used in this way, our style can be verbose, especially about object instantiation.


%
%Object interfaces do not require changes to the Java runtime or compiler,
%and they also do not introduce any new syntax. All three features of object interfaces are
%achieved by reinterpreting existing Java syntax, and are translated
%into regular Java code without loss of type-safety. Since no new
%syntax is introduced, it would be incorrect to call object interfaces
%a language extension or syntactic sugar. So we use the term
%\emph{language tuning} instead. Language tuning sits in between a
%lightweight language extension and a glorified library. Language
%tuning can offer many features usually implemented by a real language
%extension, but because it does not modify the language syntax
%pre-existing tools can work transparently on the tuned language.  To
%exploit the full benefits of language tuning,



%To formalize object interfaces, we propose Classless Java (CJ): a
%FeatherweightJava-style~\cite{Igarashi01FJ} calculus, which captures the essence
%of interfaces with default methods. The semantics of object interfaces
%is given as a syntax-directed translation from CJ to itself. In the
%resulting CJ code, all object interfaces are translated into regular CJ
%(and Java) interfaces with default methods. The translation is proved
%to be type-safe, ensuring that the translation does not
%introduce type-errors in client code.
%CJ's usefulness goes beyond serving as a
%calculus to formalize object interfaces. During the development
%process of CJ, we encountered a bug in the implementation of default
%methods for the Eclipse Compiler for Java (ECJ). For the program revealing the
%bug, ECJ behaves differently from both our formalization and Oracle's
%Java compiler.

\begin{comment}
To evaluate the usefulness of object interfaces, we illustrate
\numOfCaseStudies \bruno{needs updates}\yanlin{updated}
applications. The first application is a simple
solution to the Expression Problem~\cite{wadler98expression}, supporting independent
extensibility~\cite{zenger05independentlyextensible}, and without boilerplate code. The second
application shows how embedded DSLs using fluent interfaces~\cite{fowler2005fluentinterface}
can be easily defined using object interfaces. The third
application is a case study for a simple Maze game implemented with
multiple inheritance. For this application we show that there is a
significant reduction in the number of lines of code when compared
to an existing implementation~\cite{bono14} using plain Java 8. The last
application is a relatively larger case study of refactoring of a simple interpreter, showing that our
approach can benefit modularity and scale to real code base. Noteworthy, the first three applications are implemented
without defining a single class! The last one contains several classes that cannot be
converted to interfaces due to the limitation of our implementation of \mixin.
\end{comment}

\begin{comment}
While the Java embedding has obvious advantages from the practical
point-of-view, it also imposes some limitations that a new IB language
would not have. Supporting proper encapsulation is
difficult in Java due to limitations of Java interfaces. In particular
in Java interfaces the visibility of all methods is public. Thus
modelling private state is difficult in current Java 8. However,
 using existing design patterns~\cite{BettiniDSS13} we can emulate hiding
methods from interfaces. Furthermore Java 9 will allow private methods
in interfaces~\cite{srikanth16private}.
\end{comment}

%We discuss
%these limitations, possible workarounds, and native language support in Section~\ref{sec:futurework}.

In summary, the contributions of this paper are:
\begin{itemize}

\item {\bf IB and Object Interfaces:} which enable
  powerful programming idioms using multiple-inheritance,
  type-refinement and abstract state operations.

\item {\bf Classless Java:} a practical realization of IB in
  Java. Classless Java is implemented using annotation processing,
  allowing most tools to work transparently with our approach.
  Existing Java projects can use our approach and still be
  backward compatible with their clients, in a way that is specified
  by our safety properties.

\item {\bf Type-safe covariant mutable state:} we show how the
 combination of abstract state operations and type-refinement enables
 a form of mutable state that can be covariantly refined in a type-safe way.

%\item{\bf Type preservation guarantees:}\bruno{needs rephrasing}
%We discuss our formalization of a subset of Java8 type system and how we use this
%to characterize safety properties about our annotations.
%\item {\bf Classless Java (CJ):} A simple formal calculus that models
%the essential features of Java 8 interfaces with default methods, and
%can be used to formally define the translation of object interfaces.
%We prove several properties of the translation\footnote{Proofs and prototype implementation are available in
%  the supplementary materials.}.

\item {\bf Applications and case studies:} we illustrate the usefulness of IB
  through various examples and case studies\footnote{\url{https://github.com/YanlinWang/classless-java}}. 
  An extended version with a formal translation to Java can be found in the 
  companion technical report~\cite{classless}.

%\item{\bf Language Tuning:} We identify the concept of language tuning
%and describe object interfaces as an example. We also discuss
%how other existing approaches, such as the annotations in project
%Lombok~\cite{lombok},  can be viewed as language tuning.

\end{itemize}
